[Re] A neural model of the saccade generator in the reticular formation

### 2018-04-16 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
R1-GC-1,2,3-TEXT-1,2: made adjustments based on general comments, added explanation subscripts, corrected typos

    \usepackage{authblk}
                        \author[1, 2]{Mario Senden}
                    \author[3]{Jannis Schuecker,}
                    \author[4]{Jan Hahne,}
                    \author[3, 5, 6]{Markus Diesmann,}
                    \author[1, 2, 7]{Rainer Goebel,}
                            \affil[1]{Department of Cognitive Neuroscience, Faculty of Psychology and
Neuroscience, Maastricht University, 6201BC Maastricht, The Netherlands}
                    \affil[2]{Maastricht Brain Imaging Centre, Faculty of Psychology and Neuroscience,
Maastricht University, P.O. Box 616, 6200 MD Maastricht, The Netherlands}
                    \affil[3]{nstitute of Neuroscience and Medicine (INM-6) and Institute for Advanced
Simulation (IAS-6) and JARA Institute Brain Structure-Function
Relationships (INM-10), Jülich Research Centre, Jülich, Germany}
                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universit"at
Wuppertal, Wuppertal, Germany}
                    \affil[5]{Department of Psychiatry, Psychotherapy and Psychosomatics, Medical
Faculty, RWTH Aachen University, 52062 Aachen, Germany}
                    \affil[6]{Department of Physics, Faculty 1, RWTH Aachen University, 52062 Aachen,
Germany}
                    \affil[7]{Department of Neuroimaging and Neuromodeling, Netherlands Institute for
Neuroscience, an Institute of the Royal Netherlands Academy of Arts and
Sciences (KNAW), 1105BA Amsterdam, The Netherlands}

@@

In 2013 the European Comission launched the Human Brain Project
(\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
research infrastructure spurring pan-european collaboration.
Collaboration at such a scale, involving both empirically-oriented and
theoretical neuroscientists, offers the unique opportunity to develop
large-scale models of the brain. Specifically, individual research
groups might develop models of distinct brain regions which can
subsequently be combined into a unified whole. To facilitate this
colaboration, the HBP encourages the utilization of publicly available,
widely used, and actively developed neural simulation frameworks such as
NEST \autocite{Gewaltig2007}. In light of this, the NEST framework has
recently been extended to support simulation of functionally inspired
rate neuron models in addition to biologically grounded spiking neuron
models \autocite{Hahne2017}. Here we make use of this new functionality
of the NEST framework (v2.16.0) to provide an implementation of the
saccade generator (SG); a rate neuron model of neural circuitry in the
reticular formation proposed by Gancarz \& Grossberg
\autocite{Gancarz1998}. The SG is an integral part of the eye movement
system \autocite{Grossberg2012} and as such vital for developing
large-scale architectures of visuo-motor integration. We show that the
model translates well to the NEST framework as our implementation
faithfully reproduces all simulation results reported in the original
publication. Our code uses the Python interface \autocite{Eppler2008}
for legibility with both model and analysis scripts being implemented
using Python 2.7.12.

@@

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. While this constitutes the core SG model, the superior
collicuus (SC) is relevant for some simulations. In what follows we
briefly describe the dynamics of neurons within the horizontal component
(equivalent descriptions apply to the vertical component).

@@

Each long-lead burst neuron (\(L\)) receives excitatory external input
\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\)):

\begin{equation}
\begin{array}{ll}
\tau\frac{dL_l}{dt} &= -1.3L_l+I_l-2B_l \\\\
\tau\frac{dL_r}{dt} &= -1.3L_r+I_r-2B_r \\
\end{array}
\label{eq:llbn}\end{equation}

@@

Each excitatory burst neuron (\(E\)) receives excitatory input from the
ipsilateral LLBN, a constant arousal signal equal to \(1\), and
inhibitory input from the omnipause neuron (\(P\)):

\begin{equation}
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5L_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5L_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
given by

@@

Each inhibitory burst neuron (\(B\)) receives excitatory input from the
ipsilateral EBN:

\begin{equation}
\begin{array}{ll}
\tau\frac{dB_l}{dt} &= -2.4B_l+3E_l \\\\
\tau\frac{dB_r}{dt} &= -2.4B_r+3E_r \\
\end{array}
\label{eq:ibn}\end{equation}

@@

The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs:

\begin{equation}
\tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
\label{eq:opn}\end{equation} where \(J\) represents external electrical
stimulation and \(g(\cdot)\) again given by equation \ref{eq:gain}.

@@

Each tonic neuron (\(T\)) receives excitatory input from the ipsilateral
EBN and inhibitory input from the contralateral EBN:

\begin{equation}
\begin{array}{ll}
\tau\frac{dT_l}{dt} &= 0.1(E_l-E_r) \\\\
\tau\frac{dT_r}{dt} &= 0.1(E_r-E_l) \textrm{.} \\
\end{array}
\label{eq:tn}\end{equation} Horizontal eye position (\(\theta\)) depends
on activity of the right TN; i.e. \(\theta = 260(T_r-0.5)\).

@@

The superior colliculus (\(A\)) receives electrical stimulation F

\begin{equation}
\tau \frac{dA}{dt} =-A+F(t) \textrm{.}
\label{eq:sc}\end{equation}

@@

For a range of simulations, external input to the saccade generator is
obtained by applying a nonlinearity \(f(\cdot)\) given by
\begin{equation}
f(x) = 
\left\{
 \begin{array}{lll}
    0 \quad &\textrm{if} \quad x<0 \\
    x \quad &\textrm{if} \quad 0<x<1 \\
    1 \quad &\textrm{if} \quad x \geq 1 \\
  \end{array}
\right.\
\label{eq:pw}\end{equation} to the activity \(A\) of the SC and scaling
the result by a weight \(W\) such that

\begin{equation}
I = Wf(A)\textrm{.}
\label{eq:stim}\end{equation}

In implementing this model, we largely followed the descriptions
provided in the original publication with a number of well-motivated
exceptions. Specifically, the original model description has two
features which cannot be straightforwardly translated to NEST. First, a
nonlinear gain function is applied to a subset of inputs to EBNs and the
OPN while a linear gain function is applied to their remaining inputs.
Since NEST only applies a single gain function per neuron to each of its
inputs, we opted for using a linear gain function for EBNs and the OPN.
Furthermore, we passed those inputs requiring an additional nonlinear
gain function through an auxiliary unit instantaneously applying the
desired nonlinearity before passing the result on to EBNs and the OPN.
Second, constant input to a neuron was not hard-coded but rather
provided by an appropriately weighted bias node. Neither of these
changes lead to discrepancies with original results.

In all simulations we used the Exponential Euler (EE) method for
numerical integration of all except of tonic neurons
\autocite{Hahne2017} at a time step of \(0.05\,\mathrm{ms}\) and a time
constant of \(50\,\mathrm{ms}\). Tonic neurons were integrated using the
Euler-Maruyama (EM) method since they do not exhibit passive decay
leading to division-by-zero when using the EE method. It should be noted
that initial firing rates of neurons were not at resting equilibrium. In
order to address this, the model was allowed to evolve for
\(100\,\mathrm{ms}\), to relax towards equilibrium before applying any
input. Furthermore, as in the original publication, all activations were
bounded from below at zero. Finally, we always simulated the full model;
i.e.~both its horizontal and vertical components even if input was
applied only to one of the two.

@@

All results from the original publication have been implemented. Our
results accord very well with those reported by Gancarz \& Grossberg
\autocite{Gancarz1998}. Only a single discrepancy was observed which we
examined in depth.

The first simulation showcases the evolution of activity for each neuron
type in the horizontal SG to a constant input applied to the left LLBN.
The original publication does not report exact activation values
observed for each neuron rendering a quantitative analysis of the
accuracy of our replication impossible. However, qualitatively
activation profiles shown in figure 3 in the original publication and
those shown in figure \ref{fig:fig_1} show good correspondence. In both
implementations, the left LLBN showed a prelude of activity with left
EBN bursts beginning after the onset of LLBN activity. The right EBN
produced a small burst at the end of a saccade. Furthermore, increases
in activity of the left TN were mirrored by decreases in the right.
Finally, OPN activity dropped to zero during production of a saccade.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.45000\textwidth]{figures/fig1.eps}
\caption{\textbf{Activity profiles in the left (A) and right (B) SG.}
All activities are in response to constant input (\(\mathrm{I=1}\))
applied to the left LLBN for \(265\,\mathrm{ms}\).\label{fig:fig_1}}
\end{figure}

The second simulation shows the relation between input strength and
burst amplitude for LLBNs and EBNs. With increasing input strength, both
amplitude and duration of activation increased in long-lead and
excitatory burst neurons. As before, these results accord well with
those shown in figure 5 of the original publication.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.36000\textwidth]{figures/fig2.eps}
\caption{\textbf{Activity profiles in LLBN (A) and EBN (B).} Increased
input strength resulted in larger LLBN and EBN burst size with inputs
equal to 1 (blue), 1.75 (green), and 2.5 (red) each applied to the left
LLBN for \(85\,\mathrm{ms}\).\label{fig:fig_2}}
\end{figure}

The third simulation generates saccades in response to different input
strengths applied to the horizontal and vertical SG (figure
\ref{fig:fig_3}). As in the original publication, saccades were
generally straight with a slight tendency to curve. We extracted
quantitative estimates of saccade amplitude for the original publication
from their figure 6 and calculated the root-mean-squared error (RMSE)
between these estimates and saccade amplitudes produced by our model.
The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
0.22\textdegree, 0.14\textdegree~for the blue, green, red, turqoiuse,
and purple curves, respectively. Given that saccade amplitudes are
between 10\textdegree~and 15\textdegree, these RMSE values indicate a
close match between the two implementations.

\begin{figure}
\centering
\includegraphics[width=0.48750\textwidth,height=0.48750\textwidth]{figures/fig3.eps}
\caption{\textbf{Oblique saccades.} Inputs to the right and upward LLBNs
were \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.22}\) (green);
\(\mathrm{{I}_{r}=0.74}\) \& \(\mathrm{{I}_{u}=0.4}\) (red);
\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turqoiuse); and
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and were
applied for \(75\,\mathrm{ms}\). As expected, saccades were fairly
straight with a slight tendency to curve.\label{fig:fig_3}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.33000\textwidth,height=0.45000\textwidth]{figures/fig4.eps}
\caption{\textbf{Saccadic staircase.} Three saccades in a staircase
continued in the same direction as the initial saccade. Inputs to the
right and upward LLBNs were equal to \(\mathrm{{I}_{r}=0.2}\) and
\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). Eye position was
sampled every \(2\,\mathrm{ms}\).\label{fig:fig_4}}
\end{figure}

In the fifth simulation the average activity of the left EBN was
obtained for a series of saccades with different directions. Figure
\ref{fig:fig_5} shows a polar plot of average activity corresponding to
each saccade reflecting the neuron's tuning curve. The tuning curve we
observed for the left EBN exhibits a cardioid-like shape as was the case
in the original publication (figure 8).

\begin{longtable}[]{@{}lllllllllllllll@{}}
\caption{Direction specific inputs to SG to produce EBN tuning curve.
\label{tbl:table_1}}\tabularnewline
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endfirsthead
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endhead
\(\mathrm{{I}_{l}}\) & .00 & .00 & .00 & .00 & .20 & .45 & .63 & .70 &
.63 & .45 & .20 & .00 & .00 & .00\tabularnewline
\(\mathrm{{I}_{r}}\) & .70 & .45 & .20 & .00 & .00 & .00 & .00 & .00 &
.00 & .00 & .00 & .00 & .20 & .45\tabularnewline
\(\mathrm{{I}_{d}}\) & .00 & .00 & .00 & .00 & .00 & .00 & .00 & .00 &
.20 & .45 & .63 & .70 & .63 & .45\tabularnewline
\(\mathrm{{I}_{u}}\) & .00 & .45 & .63 & .70 & .63 & .45 & .20 & .00 &
.00 & .00 & .00 & .00 & .00 & .00\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=0.37500\textwidth,height=0.37500\textwidth]{figures/fig5.eps}
\caption{\textbf{EBN tuning curve.} Tuning curve of the left EBN
exhibiting a cardioid-like shape. Inputs to the SG producing the desired
saccades can be found in table \ref{tbl:table_1}. Each of these inputs
was applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
\end{figure}

\pagebreak
The sixth simulation reported in the original publication was designed
to replicate results of Stanford \textit{et al.}
\autocite{Stanford1996}. These authors stimulated the superior
colliculus (SC) at various frequencies and measured the resulting
saccade amplitude, duration, and velocity; showing that ampltidude
saturates before velocity. Our implementation of the SG model was
capable of replicating these results. However, reproducing simulation
results reported by Gancarz \& Grossberg \autocite{Gancarz1998} with our
implementation was complicated by the fact that the stimulation protocol
given by the authors lead to the production of two rather than a single
saccade for larger stimulation intensities. Furthermore, the authors did
not report their criteria for identifying saccade on- and offsets.
Stanford \textit{et al.} \autocite{Stanford1996} used velocity criteria
to determine onset (\(v\textgreater30\,\mathrm{deg/s}\)) and offset
(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provided us
with explicit criteria, velocity did not always drop below
\(30\,\mathrm{deg/s}\) after the first saccade before rising again with
the second. To determine the offset of a saccade in those cases, we
found the local minimum between the end of the first and the beginning
of the second saccade. With these criteria in place, we stimulated the
SC. Results of our simulation are shown in figure \ref{fig:fig_6}. We
extracted quantitative results of the original publication from their
figure 9 and calculated the RMSE between these estimates and our results
for each curve. The RMSE was 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
\(4.45\,\mathrm{deg/s}\) for saccade amplitude, duration, and peak
velocity, respectively. Our results accord thus very well with those
reported in the original publication. This includes the observation that
saccade amplitude and especially saccade duration were largest for a
stimulation intensity of 1.2. It is conceivable that this intensity
marks the point after which the SG generator starts producing two rather
than a single saccade with the eye movement at this intensity merging
two saccades and thus being stretched out.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.37500\textwidth]{figures/fig6.eps}
\caption{\textbf{Effect of stimulation frequency on saccade (A)
amplitude, (B) duration, and (C) velocity.} Saccade amplitude (A)
saturated before saccade velocity (C). A range of unitless values
varying between 1 and 2.4 at increments of 0.2 were applied to the SC
(reflecting stimulation at different frequencies). The connection weight
from SC to LLBN was \(\mathrm{W=2}\) and stimulation duration was
\(125\,\mathrm{ms}\).\label{fig:fig_6}}
\end{figure}

>>>>

    \usepackage{authblk}
                        \author[1, 2]{Mario Senden}
                    \author[3]{Jannis Schuecker,}
                    \author[4]{Jan Hahne,}
                    \author[3, 5, 6]{Markus Diesmann,}
                    \author[1, 2, 7]{Rainer Goebel,}
                            \affil[1]{Department of Cognitive Neuroscience, Faculty of Psychology and
Neuroscience, Maastricht University, 6201BC Maastricht, The Netherlands}
                    \affil[2]{Maastricht Brain Imaging Centre, Faculty of Psychology and Neuroscience,
Maastricht University, P.O. Box 616, 6200 MD Maastricht, The Netherlands}
                    \affil[3]{Institute of Neuroscience and Medicine (INM-6) and Institute for
Advanced Simulation (IAS-6) and JARA Institute Brain Structure-Function
Relationships (INM-10), Jülich Research Centre, Jülich, Germany}
                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universität
Wuppertal, Wuppertal, Germany}
                    \affil[5]{Department of Psychiatry, Psychotherapy and Psychosomatics, Medical
Faculty, RWTH Aachen University, 52062 Aachen, Germany}
                    \affil[6]{Department of Physics, Faculty 1, RWTH Aachen University, 52062 Aachen,
Germany}
                    \affil[7]{Department of Neuroimaging and Neuromodeling, Netherlands Institute for
Neuroscience, an Institute of the Royal Netherlands Academy of Arts and
Sciences (KNAW), 1105BA Amsterdam, The Netherlands}

@@

In 2013 the European Commission launched the Human Brain Project
(\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
research infrastructure spurring pan-European collaboration.
Collaboration at such a scale, involving both empirically-oriented and
theoretical neuroscientists, offers the unique opportunity to develop
large-scale models of the brain. Specifically, individual research
groups might develop models of distinct brain regions which can
subsequently be combined into a unified whole. To facilitate this
collaboration, the HBP encourages the utilization of publicly available,
widely used, and actively developed neural simulation frameworks such as
NEST \autocite{Gewaltig2007}. In light of this, the NEST framework has
recently been extended to support simulation of functionally inspired
rate neuron models in addition to biologically grounded spiking neuron
models \autocite{Hahne2017}. Here we make use of this new functionality
of the NEST framework (2.16.0) to provide an implementation of the
saccade generator (SG); a rate neuron model of neural circuitry in the
reticular formation proposed by Gancarz \& Grossberg
\autocite{Gancarz1998}. The SG is an integral part of the eye movement
system \autocite{Grossberg2012} and as such vital for developing
large-scale architectures of visuo-motor integration. We show that the
model translates well to the NEST framework as our implementation
faithfully reproduces all simulation results reported in the original
publication. Our code uses the Python interface \autocite{Eppler2008}
for legibility with both model and analysis scripts being implemented
using Python 2.7.12. However, the code is compatible with Python 3 as
well (tested with version 3.5.2).

@@

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. While this constitutes the core SG model, the superior
colliculus (SC) is relevant for some simulations. In what follows we
briefly describe the dynamics of neurons within the horizontal component
(equivalent descriptions apply to the vertical component). Please note
that subscripts \(\mathrm{l}\), \(\mathrm{r}\), \(\mathrm{d}\) and
\(\mathrm{u}\) indicate whether a neuron controls upward, downward,
leftward or rightward saccade direction, respectively.

@@

Each long-lead burst neuron (\(L\)) receives excitatory external input
\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\))

\begin{equation}
\begin{array}{ll}
\tau\dot L_\mathrm{l} &= -1.3L_\mathrm{l}+I_\mathrm{l}-2B_\mathrm{l} \\\\
\tau\dot L_\mathrm{r} &= -1.3L_\mathrm{r}+I_\mathrm{r}-2B_\mathrm{r} \textrm{.} \\
\end{array}
\label{eq:llbn}\end{equation}

@@

Each excitatory burst neuron (\(E\)) receives excitatory input from the
ipsilateral LLBN, a constant arousal signal equal to \(1\), and
inhibitory input from the contralateral LLBN and the omnipause neuron
(\(P\))

\begin{equation}
\begin{array}{ll}
\tau\dot E_\mathrm{l} &= -3.5E_\mathrm{l}+(2-E_\mathrm{l})(5L_\mathrm{l}+1)-(E_\mathrm{l}+1)(10L_\mathrm{r}+20g(P)) \\\\
\tau\dot E_\mathrm{r} &= -3.5E_\mathrm{r}+(2-E_\mathrm{r})(5L_\mathrm{r}+1)-(E_\mathrm{r}+1)(10L_\mathrm{l}+20g(P)) \textrm{,} \\
\end{array}
\label{eq:ebn}\end{equation} where the nonlinear gain function
\(g(\cdot)\) is given by

@@

Each inhibitory burst neuron (\(B\)) receives excitatory input from the
ipsilateral EBN

\begin{equation}
\begin{array}{ll}
\tau\dot B_\mathrm{l} &= -2.4B_\mathrm{l}+3E_\mathrm{l} \\\\
\tau\dot B_\mathrm{r} &= -2.4B_\mathrm{r}+3E_\mathrm{r} \textrm{.} \\
\end{array}
\label{eq:ibn}\end{equation}

@@

The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs

\begin{equation}
\tau\dot P = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_\mathrm{l})+g(L_\mathrm{r})+g(L_\mathrm{u})+g(L_\mathrm{d})) \textrm{,}
\label{eq:opn}\end{equation} where \(J\) represents external electrical
stimulation and \(g(\cdot)\) is again given by equation \ref{eq:gain}.

@@

Each tonic neuron (\(T\)) receives excitatory input from the ipsilateral
EBN and inhibitory input from the contralateral EBN

\begin{equation}
\begin{array}{ll}
\tau\dot T_\mathrm{l} &= 0.1(E_\mathrm{l}-E_\mathrm{r}) \\\\
\tau\dot T_\mathrm{r} &= 0.1(E_\mathrm{r}-E_\mathrm{l}) \textrm{.} \\
\end{array}
\label{eq:tn}\end{equation} The horizontal eye position (\(\theta\))
depends on the activity of the right TN; i.e.
\(\theta = 260(T_\mathrm{r}-0.5)\).

@@

The superior colliculus (\(A\)) receives electrical stimulation \(F\)
and evolves according to

\begin{equation}
\tau \dot A =-A+F \textrm{.}
\label{eq:sc}\end{equation}

@@

For a range of simulations, external input to the saccade generator is
obtained by applying a nonlinearity \(f(\cdot)\) given by
\begin{equation}
f(x) = 
\left\{
 \begin{array}{lll}
    0 \quad &\textrm{if} \quad x \leq 0 \\
    x \quad &\textrm{if} \quad 0<x<1 \\
    1 \quad &\textrm{if} \quad x \geq 1 \\
  \end{array}
\right.\
\label{eq:pw}\end{equation} to the activity \(A\) of the SC and scaling
the result by a weight \(W\) such that

\begin{equation}
I = Wf(A) \textrm{.}
\label{eq:stim}\end{equation}

In implementing this model, we largely follow the descriptions provided
in the original publication with a number of motivated exceptions.
Specifically, the original model description has two features which
cannot be straightforwardly translated to NEST. First, a nonlinear gain
function is applied to a subset of inputs to EBNs and the OPN while a
linear gain function is applied to their remaining inputs. Since NEST
applies the same gain function to all inputs of a neuron, we opt for
using a linear gain function for EBNs and the OPN, but pass those inputs
requiring an additional nonlinear gain function through an auxiliary
unit instantaneously applying the desired nonlinearity before passing
the result on to EBNs and the OPN. Second, constant input to a neuron is
not hard-coded but rather provided by an appropriately weighted bias
node. Neither of these changes leads to discrepancies with original
results.

In all simulations we use the Exponential Euler (EE) method for
numerical integration of all except tonic neurons \autocite{Hahne2017}
at a time step of \(0.05\,\mathrm{ms}\) and a time constant of
\(50\,\mathrm{ms}\). Tonic neurons are integrated using the forward
Euler method since they do not exhibit passive decay leading to
division-by-zero when using the EE method. It should be noted that the
initial firing rates of neurons are not in equilibrium. In order to
address this, the model evolves for \(100\,\mathrm{ms}\) to relax
towards equilibrium before any input is applied. Furthermore, as in the
original publication, any activity is bounded from below at zero.
Finally, we always simulate the full model; i.e.~both its horizontal and
vertical components even if input is applied only to one of the two.

@@

All simulations of the original publication are repeated. Our results
accord very well with those reported by Gancarz \& Grossberg
\autocite{Gancarz1998}. Only a single discrepancy is observed which we
examine in depth.

The first simulation showcases the evolution of activity for each neuron
type in the horizontal SG for a constant input applied to the left LLBN.
The original publication does not report the exact activation values
observed for each neuron rendering a quantitative analysis of the
accuracy of our replication impossible. However, qualitatively
activation profiles shown in figure 3 in the original publication and
those shown in figure \ref{fig:fig_1} show good correspondence. In both
implementations, the left LLBN exhibits a prelude of activity with left
EBN bursts beginning after the onset of LLBN activity. The right EBN
produces a small burst at the end of a saccade. Furthermore, increases
in activity of the left TN are mirrored by decreases in the right.
Finally, OPN activity drops to zero during production of a saccade.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.45000\textwidth]{figures/fig1.eps}
\caption{\textbf{Activity profiles in the left (A) and right (B) SG.}
All activities are in response to constant input (\(\mathrm{I=1}\))
applied to the left LLBN for \(265\,\mathrm{ms}\).\label{fig:fig_1}}
\end{figure}

The second simulation shows the relation between input strength and
burst amplitude for LLBNs and EBNs. With increasing input strength, both
amplitude and duration of activation increase in long-lead and
excitatory burst neurons. As before, these results accord well with
those shown in figure 5 of the original publication.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.36000\textwidth]{figures/fig2.eps}
\caption{\textbf{Activity profiles in LLBN (A) and EBN (B).} Increased
input strength resulted in larger LLBN and EBN burst size with inputs
equal to 1 (blue), 1.75 (green), and 2.5 (red) each applied to the left
LLBN for \(85\,\mathrm{ms}\).\label{fig:fig_2}}
\end{figure}

The third simulation generates saccades in response to different input
strengths applied to the horizontal and vertical SG (figure
\ref{fig:fig_3}). As in the original publication, saccades are generally
straight with a slight tendency to curve. Using
\href{http://plotdigitizer.sourceforge.net/}{PlotDigitizer}, we extract
saccade endpoints displayed in figure 6 of the original publication.
This provides us with estimates of saccade amplitude allowing us to
quantitatively asses our results in terms of the root-mean-squared error
(RMSE). The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
0.22\textdegree, 0.14\textdegree~for the blue, green, red, turquoise,
and purple curves, respectively. Given that saccade amplitudes are
between 10\textdegree~and 15\textdegree, these RMSE values indicate a
close match between the two implementations.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.48750\textwidth,height=0.48750\textwidth]{figures/fig3.eps}
\caption{\textbf{Oblique saccades.} Inputs to the right and upward LLBNs
are \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.22}\) (green);
\(\mathrm{{I}_{r}=0.74}\) \& \(\mathrm{{I}_{u}=0.4}\) (red);
\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turquoise); and
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and are
applied for \(75\,\mathrm{ms}\).\label{fig:fig_3}}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.33000\textwidth,height=0.45000\textwidth]{figures/fig4.eps}
\caption{\textbf{Saccades in a staircase.} Eye positions exhibited in a
saccade staircase simulation. Dense clusters reflect saccade endpoints.
A total of three saccades are exhibited with subsequent saccades
continuing in the same direction as the initial saccade. Inputs to the
right and upward LLBNs are equal to \(\mathrm{{I}_{r}=0.2}\) and
\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). The eye position is
sampled every \(2\,\mathrm{ms}\).\label{fig:fig_4}}
\end{figure}

In the fifth simulation the average activity of the left EBN is obtained
for a series of saccades with different directions. Figure
\ref{fig:fig_5} shows a polar plot of average activity corresponding to
each saccade. This is the neuron's tuning curve. The tuning curve we
observe for the left EBN exhibits a cardioid-like shape as in the
original publication (figure 8).

\pagebreak

\begin{longtable}[c]{@{}lllllllllllllll@{}}
\caption{Direction specific inputs to SG to produce EBN tuning curve.
\label{tbl:table_1}}\tabularnewline
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endfirsthead
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endhead
\(\mathrm{{I}_{l}}\) & .00 & .00 & .00 & .00 & .20 & .45 & .63 & .70 &
.63 & .45 & .20 & .00 & .00 & .00\tabularnewline
\(\mathrm{{I}_{r}}\) & .70 & .45 & .20 & .00 & .00 & .00 & .00 & .00 &
.00 & .00 & .00 & .00 & .20 & .45\tabularnewline
\(\mathrm{{I}_{d}}\) & .00 & .00 & .00 & .00 & .00 & .00 & .00 & .00 &
.20 & .45 & .63 & .70 & .63 & .45\tabularnewline
\(\mathrm{{I}_{u}}\) & .00 & .45 & .63 & .70 & .63 & .45 & .20 & .00 &
.00 & .00 & .00 & .00 & .00 & .00\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.37500\textwidth,height=0.37500\textwidth]{figures/fig5.eps}
\caption{\textbf{EBN tuning curve.} Tuning curve of the left EBN
exhibiting a cardioid-like shape. Inputs to the SG producing the
saccades are given in table \ref{tbl:table_1}. Each of these inputs is
applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
\end{figure}

The sixth simulation reported in the original publication is designed to
replicate results of Stanford \textit{et al.} \autocite{Stanford1996}.
These authors stimulated the superior colliculus (SC) at various
frequencies and measured the resulting saccade amplitude, duration, and
velocity; showing that amplitude saturates before velocity. Our
implementation of the SG model is capable of replicating these results.
However, reproducing simulation results reported by Gancarz \& Grossberg
\autocite{Gancarz1998} with our implementation is complicated by the
fact that the stimulation protocol given by the authors leads to the
production of two rather than a single saccade for larger stimulation
intensities. Furthermore, the authors do not report their criteria for
identifying saccade on- and offsets. Stanford \textit{et al.}
\autocite{Stanford1996} use velocity criteria to determine onset
(\(v\textgreater30\,\mathrm{deg/s}\)) and offset
(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provides us
with explicit criteria, velocity does not always drop below
\(30\,\mathrm{deg/s}\) after the first saccade before rising again with
the second. To determine the offset of a saccade in those cases, we find
the local minimum between the end of the first and the beginning of the
second saccade. With these criteria in place, we stimulate the SC.
Results of our simulation are shown in figure \ref{fig:fig_6}. Again we
use PlotDigitizer to extract a vector of saccade amplitude, duration and
peak velocity observed at each stimulation frequency in the original
publication from their figure 9 and calculate the RMSE between these and
our curves. The RMSE is 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
\(4.45\,\mathrm{deg/s}\) for saccade amplitude, duration, and peak
velocity, respectively. Our results accord thus very well with those
reported in the original publication. This includes the observation that
saccade amplitude and especially saccade duration are largest for a
stimulation intensity of 1.2. It is conceivable that this intensity
marks the point after which the SG starts producing two rather than a
single saccade with the eye movement at this intensity merging two
saccades and thus being stretched out.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.37500\textwidth]{figures/fig6.eps}
\caption{\textbf{Effect of stimulation frequency on saccade amplitude
(A), duration (B), and velocity (C).} A range of unitless values varying
between 1 and 2.4 at increments of 0.2, reflecting stimulation at
different frequencies ,is applied to the SC. The connection weight from
SC to LLBN is \(\mathrm{W=2}\) and stimulation duration is
\(125\,\mathrm{ms}\).\label{fig:fig_6}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth,height=0.90000\textwidth]{figures/fig7.eps}
\caption{\textbf{Trading saccade velocity for duration.} Duration and
velocity of a saccade can be traded while keeping amplitude constant. To
produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
publication; solid curve). To produce a low velocity saccade an input of
\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\) (dashed curve). Given the
shape of the input curve produced by our implementation of the original
equations (A), a second saccade can be observed for both high and low
velocity saccades. If the input curves reported in the original
manuscript are recreated in an ad-hoc fashion (B), the responses of all
model neurons matches those in \autocite{Gancarz1998}\label{fig:fig_7}}
\end{figure}
